<!--
  %\VignetteEngine{knitr}
  %\VignetteIndexEntry{Importing GSEA}
  %\VignetteDepends{categoryComparePaper}
-->

# Importing GSEA Data

How would we work with GSEA data? We are going to assume that we are working with the simple text results (what GSEA reports as .xls) from GSEA. We are using the results from the same comparison performed in the `EnrichmentMap` tutorial for GSEA. We should note that the results files generated were generated by R.M. Flight on Feb 17, 2014, but the input files were downloaded from the EnrichmentMap website.

```{r locData}
fileLoc <- system.file("extdata", package="categoryComparePaper")
```

## Import GMT

We first need to be able to import GSEA's `gmt` files that define the gene sets.

```{r draftGMT}
gmtfile <- file.path(fileLoc, "Human_GO_AllPathways_no_GO_iea_April_15_2013_symbol.gmt.gz")
readGMT <- function(gmtfile){
  gmtContents <- scan(gmtfile, what="char", sep="\n", encoding="latin1")
  nSet <- length(gmtContents)
  gmtID <- character(nSet)
  gmtDesc <- character(nSet)
  
  gmtSplit <- strsplit(gmtContents, "\t", fixed=F)
    
  gmtData <- lapply(seq(1, nSet), function(inLoc){
    gmtID[inLoc] <<- gmtSplit[[inLoc]][1]
    gmtDesc[inLoc] <<- gmtSplit[[inLoc]][2]
    nSplit <- length(gmtSplit[[inLoc]])
    unique(gmtSplit[[inLoc]][seq(3, nSplit)])
  })
  names(gmtData) <- gmtID
  return(list(id=gmtID, description=gmtDesc, data=gmtData))
}
gmtData <- readGMT(gmtfile)
```

## Import Results

```{r readResults}
fileList <- list(ES12.Down="gsea_nt12_results.xls",
                 ES12.Up="gsea_es12_results.xls",
                 ES24.Down="gsea_nt24_results.xls",
                 ES24.Up="gsea_es24_results.xls")
fileList <- lapply(fileList, function(x){file.path(fileLoc, x)})

readGSEAResults <- function(fileList, pUse="NOM p-val", pCut=0.01){
  pUse <- make.names(pUse)
  outResults <- lapply(fileList, function(x){
    tmpRes <- read.table(x, header=TRUE, sep="\t", quote="", stringsAsFactors=FALSE)
    sigLoc <- which(names(tmpRes) %in% pUse)
    sigEntry <- tmpRes[,sigLoc] <= pCut
    return(new("ccSigList", sigID=tmpRes[sigEntry, 'NAME']))
  })
  return(outResults)
}

gseaRes <- readGSEAResults(fileList)
```

## Run categoryCompare

```{r runCC}
geneAnnMapping <- new("namedList", .Data=gmtData$data, names=names(gmtData$data))
ccGSEA <- new("GENccEnrichResult", gseaRes, categoryName="gsea", geneAnnMapping=geneAnnMapping, overlapType="overlap", annDescription=gmtData$description)

gseaOpts <- new("ccOptions", listNames=names(ccGSEA), compareNames=c("ES12.Up",
                                                                     "ES12.Down",
                                                                     "ES24.Up",
                                                                     "ES24.Down",
                                                                     "ES12.Up,ES24.Up",
                                                                     "ES12.Up,ES24.Down",
                                                                     "ES12.Down,ES24.Up",
                                                                     "ES12.Down,ES24.Down"))
compGSEA <- ccCompare(ccGSEA, gseaOpts)
```

## Visualize the results

```{r visResults, eval=FALSE}
compGSEA <- breakEdges(compGSEA, 0.8)
cwGSEA <- ccOutCyt(compGSEA, gseaOpts, postText="GSEA", rpcPort=9001)
```

